---
title: "Index"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Methods 

### Chance Event Function: Generating Conditional Values

The `chance_event` function generates conditional values with: 

   \[
   occurrence =
   \begin{cases}
   rbinom(length(value\_if), 1, chance) & if  !one\_draw \\
   rbinom(1, 1, chance) & if  one\_draw
   \end{cases}
   \]
   
and final values with: 

   \[
   chance\_event = occurrence \cdot value\_if + (1 - occurrence) \cdot value\_if\_not
   \]

### Value Varier Function: Generating Synthetic Data

```{r vv}
source("functions/vv.R")
```

The `vv` function generates synthetic data with varying means and coefficients of variation with: 

- \( n \): Number of data points to generate.
- \( var\_mean \): Mean value for the data.
- \( var\_CV \): Coefficient of Variation (CV) for the data.
- \( distribution \): Distribution of the data (e.g., normal distribution).
- \( absolute\_trend \): Absolute trend to add to the data.
- \( relative\_trend \): Relative trend to apply to the data.
- \( lower\_limit \): Lower limit for generated data.
- \( upper\_limit \): Upper limit for generated data.

These are used to calculate annual means: 

   - If both absolute and relative trends are absent, the annual means remain constant:

     \[
     annual\_means = var\_mean
     \]

   - If an absolute trend is present, annual means are adjusted linearly:

     \[
     annual\_means = var\_mean + absolute\_trend \cdot (0, 1, \ldots, n-1)
     \]

   - If only a relative trend is present, annual means change multiplicatively:

     \[
     annual\_means = var\_mean \cdot (1 + relative\_trend / 100)^{(0, 1, \ldots, n-1)}
     \]

   - Generate random data points based on the specified distribution using calculated annual means:

     \[
     out \sim \mathcal{N}(annual\_means, \lvert annual\_means \rvert \cdot \frac{var\_CV}{100})
     \]

   - If specified, limit data points to the defined lower and upper limits:

     \[
     out = \max(out, lower\_limit), \quad out = \min(out, upper\_limit)
     \]

### Net Present Value (NPV)

```{r NPV}
source("functions/discount.R")
```

The discounted value of a time series value \(x_t\) at time \(t\) can be calculated using the formula:

   \[
   \text{Discounted Value}_t = \frac{x_t}{\left(1 + \frac{\text{Discount Rate}}{100}\right)^{t-1}}
   \]

   Where:
   
   - \(x_t\) is the value at time \(t\),
   
   - \(\text{Discount Rate}\) is the discount rate provided to the function,
   
   - \(t\) is the time period (starting from 1).

2. **Net Present Value (NPV)** (if `calculate_NPV` is set to `TRUE`):
   The Net Present Value of a time series with discounted values can be calculated as the sum of all discounted values:

   \[
   \text{NPV} = \sum_{t=1}^{n} \frac{x_t}{\left(1 + \frac{\text{Discount Rate}}{100}\right)^{t-1}}
   \]

   Where \(n\) is the number of time periods in the time series.

## Monte Carlo Simulation Function

```{r MC}
source("functions/mcSimulation.R")
```

The `mcSimulation` function performs Monte Carlo simulations to estimate model outputs based on provided parameters and a model function.

## Monte Carlo Simulation Process

The Monte Carlo simulation process generates a set of estimated model outputs based on random input samples, providing a distribution of potential outcomes including:

- random input generation with \( x_1, x_2, \ldots, x_n \) as a set of \( n \) random input samples drawn from a given distribution, representing the input parameters of the model.

- a model function \( f(x) \) that maps the input data \( x \) to the corresponding model output \( y \). This can be expressed as:
   
   \[
   y_i = f(x_i), \quad i = 1, 2, \ldots, n
   \]

- Estimated model outputs \( \hat{y}_1, \hat{y}_2, \ldots, \hat{y}_n \) obtained by applying the model function to the random input samples:
   
   \[
   \hat{y}_i = f(x_i), \quad i = 1, 2, \ldots, n
   \]





